---
title: 'TReNA: Past, Present, and Future'
author: "Matt Richards"
date: "5/4/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is TReNA?

It's an R package for:

- Feature selection
- Footprinting
- Functional genomics

## Using TReNA

When using TReNA, you'll need an assay matrix of expression data where:

1. Rows are different genes
2. Columns are different samples

The TReNA package also comes with its own example data:

```{r}
suppressMessages(library(TReNA))
load(system.file(package="TReNA", 
                 "extdata/ampAD.154genes.mef2cTFs.278samples.RData"))
mtx.sub[1:3,1:5]
```

## Creating and Solving a TReNA Object

```{r}
trena <- TReNA(mtx.assay = mtx.sub, solver = "lasso")
target.gene <- "MEF2C"
tfs <- setdiff(rownames(mtx.sub), target.gene)
tbl.out <- solve(trena, target.gene, tfs)
```

```{r echo = FALSE}
knitr::kable(head(tbl.out,4))
```

## TReNA in December 2016

* 3 different "solvers"
     * LASSO
     * Random Forest
     * Bayes Spike
 
* Footprint Finders
     * Pull out footprints from a:
         * Gene
         * Region

## Distribution Effects

```{r}
hist(mtx.sub, main = "As-is Data")

```

## Transforming with arcsinh

```{r}
mtx.asinh <- asinh(mtx.sub)
hist(mtx.asinh, main = "Asinh-Transformed Data")

```

## Solvers effects

In addition to the distributions, solver choice can greatly affect results.

```{r echo=FALSE, cache=TRUE}
source(system.file(package="TReNA", 
                  "utils/evaluateAllSolvers.R"))
printf <- function(...) print(noquote(sprintf(...)))
suppressMessages(tbl.all <- assess_ampAD154AllSolversAndDistributions())
```

## Pair Plots

```{r echo = FALSE}
pairs(tbl.all[c(2,4,6,8,10,12)],
      labels = names(tbl.all)[c(2,4,6,8,10,12)])
```

## Correlations

```{r echo = FALSE}
knitr::kable(signif(cor(tbl.all[c(2,4,6,8,10,12)]),2))
```

## Takeaways from Past TReNA

* Distribution matters
    * As-is data is often quite skewed
    * We recommend asinh or VOOM 
* Solver choice matters
    * Are 3 solvers sufficient?
    * What else could/should we add?
    * How do we compare between solvers?
    
## We've Added 5 (6) Solvers

* [Square Root LASSO](https://cran.r-project.org/web/packages/flare/index.html)
* [Pearson Correlation](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/cor.html)
* [Spearman Correlation](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/cor.html)
* [LASSO P-Values](https://cran.r-project.org/web/packages/lassopv/index.html)
* [Ridge Regression](https://cran.r-project.org/web/packages/glmnet/index.html)
* Ensemble Solver

## The Ensemble Solver class

```{r cache = TRUE, echo = FALSE}

trena <- TReNA(mtx.assay = mtx.asinh, solver = "ensemble")
target.gene <- "MEF2C"
tfs <- setdiff(rownames(mtx.asinh), target.gene)
tbl.out <- solve(trena, target.gene, tfs)
knitr::kable(head((tbl.out[,c(1:3,6:9)])))
```

## Examining Solver PCA

```{r echo = FALSE, fig.height=5,fig.width=5}

   tbl.out <- tbl.out[,c(1:7)]
   tbl.scale <- scale(tbl.out[,-1])
   rownames(tbl.scale) <- tbl.out$gene
   
   # do pca on the matrix
   pca <- prcomp(tbl.scale, center=FALSE, scale.=FALSE)
   # view pca
   #par(mar = c(1,1,0.5,1))
   biplot(pca, c(1,2))
```

## Examining Solver PCA

```{r echo = FALSE, fig.height=5,fig.width=5}

   biplot(pca, c(1,3))
```

## The CandidateFilter Class

* Feature selection is nice, but nothing new
* Our value added is largely in how we filter predictors
    * VarianceFilter class
    * **FootprintFilter** class

## Using a VarianceFilter

```{r}
variance.filter <- VarianceFilter(mtx.assay = mtx.asinh)
tf.list <- getCandidates(variance.filter, extraArgs = list("target.gene" = "MEF2C", "var.size" = 0.5))
str(tf.list)
```

## Using a FootprintFilter

```{r echo = FALSE}
footprint.filter <- FootprintFilter(mtx.assay = mtx.asinh)
db.address <- system.file(package = "TReNA", "extdata")
genome.db.uri <- paste("sqlite:/",db.address,"genome.sub.db", sep = "/")
project.db.uri <- paste("sqlite:/",db.address,"project.sub.db", sep = "/")
target.gene <- "MEF2C"
extraArgs = list("target.gene" = target.gene,
                                                        "genome.db.uri"=genome.db.uri, 
                                                        "project.db.uri" = project.db.uri,
                                                        "size.upstream" = 1000,
                                                        "size.downstream" = 1000)
```

```{r cache=TRUE}
tfs <- getCandidates(footprint.filter, extraArgs)
str(tfs)
```

## The FootprintFinder Class

We have public databases of footprints (with more on the way) 

```{r eval = FALSE}
genome.db.uri    <- "postgres://bddsrds.globusgenomics.org/hg38"
footprint.db.uri <- "postgres://bddsrds.globusgenomics.org/brain_hint"
fpf <- FootprintFinder(genome.db.uri, footprint.db.uri, quiet=FALSE)
tbl.fp <- getFootprintsInRegion(fpf, "chr5", 88822685, 89011824)
str(tbl.fp)
```

Unfortunately, the databases decided to misbehave right now...

## Summing Up TReNA at Present

Current Workflow:

1. Start with a matrix of expression data
2. Transform the matrix to alter the distribution
3. Use a filter to wean down predictors
4. Run a feature selection with chosen solver

Where do I get TReNA?

* Accepted to Bioconductor Development Branch
* [PriceLab Github](http://trena.org)

## What's Next?

* More databases
    * 22 tissue types
    * 3 footprinting methods
    * 2 different alignments
* Structural changes to TReNA
    * More filtering options
        * DHS filter
        * GO filter
* Web application
    * Ability to build TReNA workflows

## Thank you

Special thanks to:

* Nathan Price
* Paul Shannon
* Cory Funk
* Max Robinson
* Rory Donovan-Maiye

For More on TReNA, visit the [current TReNA webpage](https://pricelab.github.io/TReNA/)
