---
title: 'TReNA: Past, Present, and Future'
author: "Matt Richards"
date: "5/4/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is TReNA?

It's an R package for:

- Feature selection
- Footprinting
- Functional genomics

## Using TReNA

When using TReNA, you'll need an assay matrix of expression data where:

1. Rows are different genes
2. Columns are different samples

The TReNA package also comes with its own example data:

```{r}
suppressMessages(library(TReNA))
load(system.file(package="TReNA", 
                 "extdata/ampAD.154genes.mef2cTFs.278samples.RData"))
mtx.sub[1:3,1:5]
```

## TReNA in December 2016

* 3 different "solvers"
     * LASSO
     * Random Forest
     * Bayes Spike
 
* Footprint Finders
     * Pull out footprints from a:
     * Gene
     * Region

## Using TReNA




## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

